### 2014-03-13 conclusion.tex
Split paper in a tex file per section to make merge issues less likely and easier to manage

In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore, we looked at the main disadvantages and problems Tor is currently facing. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

From this table, we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of attacks.

While Tor also still has a lot to improve on, further research and development is needed to improve the current decentralized systems and develop a decentralized system that matches Tor's current standards.