### 2014-03-13 systems-comparison.tex

In the last section, we discussed existing decentralized peer-to-peer networks. In this section we will make a comparison between these networks. For each network, we will explain the following features:

\begin{itemize}
	\item{Compatibility with Tor:} If the network is compatible with Tor, the network or some features of the network could be used for the decentralization of Tor.
	\item{Public implementation:} For each network we will explain whether a public implementation is available.
	\item{Used in practice:} Is the network used in practice?
	\item{Attack resistance:} We will explain the various weaknesses and possible attacks of each network.
	\item{Anonimity:} Does the network hide the identities of the sender and/or receiver?
\end{itemize}