### 2014-03-20 table_old.tex
Redesigned the table

\begin{landscape}

\begin{center}
    \begin{longtable}{ | p{1.9cm} | p{0.8cm} | p{1.6cm} | p{2.2cm} | p{2.0cm} | p{2cm} | p{2cm} | p{2cm} |}
  		\caption{A comparison of decentralized peer-to-peer overlay networks} \\ \hline
	    \textbf{Name} & \textbf{Year} & \textbf{Tor interoperability} & \textbf{Public implementation} & \textbf{Used in practice} & \textbf{Attack resistance} & \textbf{Provides anonymity} \\ \hline
		\endfirsthead
		\multicolumn{7}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\ \hline
	    \textbf{Name} & \textbf{Year} & \textbf{Tor interoperability} & \textbf{Public implementation} & \textbf{Used in practice} & \textbf{Attack resistance} & \textbf{Provides anonymity} \\ \hline
		\endhead
		\hline \multicolumn{7}{r}{\textit{Continued on next page}} \\
		\endfoot
		\hline
		\endlastfoot

		Gnutella & ~2000 & no & yes & yes, has millions of users & Vulnerable for DDoS attacks and attacks from misbehaving peers & Queries are anonymous, IP addresses are revealed when downloading files.\\ \hline
		Freenet & 2001 & no & yes, implementations are based on the FCP (Freenet Client Protocol) & yes & Vulnerable for man-in-the-middle attacks and Trojan attacks & yes\\ \hline
		Tapestry & 2001 & no & yes & yes, several applications are based on it such as OceanStore and Bayeux & Vulnerable for man-in-the-middle attacks and Trojan attacks & no\\ \hline
		Pastry & 2001 & no & yes & yes & Vulnerable for man-in-the-middle attacks and Trojan attacks & no\\ \hline
		Anonymizing peer-to-peer proxy (AP3) & 2004 & no & no & no & Provides anonymity even in the case of a large-scale attack by a coordinated set of malicious nodes & yes\\ \hline
		Tarzan & 2002 & no & no(?) & no & ? & yes\\ \hline
		MorphMix & 2002 & no & yes & no & yes, although depending on the number of colluding malicious nodes & yes\\ \hline
		I2P & 2003 & no & yes & yes & ? & yes\\ \hline
		Salsa & 2006 & no & ? & ? & ? & ?\\ \hline
		Tribler & 2008 & no, based on BitTorrent & yes & yes & ? & not yet\\ \hline
		Torks & 2009 & yes & implementation exists, but not public(?) & no & ? & yes\\ \hline
		NISAN & 2009 & no & no & no & ? & yes\\ \hline
    \end{longtable}
\end{center}

\end{landscape}