### 2014-03-20 table.tex
Added info about morphmix, salsa and tribler. Also renamed torks to torsk

\begin{center}
    \begin{longtable}{ | p{1.9cm} | p{0.8cm} | p{1.6cm} | p{2.2cm} | p{2.2cm} | p{2.5cm} | p{2.5cm} | p{2.5cm} | p{2cm} | }
  		\caption{A comparison of decentralized peer-to-peer overlay networks} \\ \hline
	    \textbf{Name} & \textbf{Year} & \textbf{Tor interoperability} & \textbf{Public implementation} & \textbf{Used in practice} & \textbf{Secure against DDoS attack} & \textbf{Secure against Sybil Attack} & \textbf{Secure against man-in-the-middle attack} & \textbf{Provides anonymity} \\ \hline
		\endfirsthead
		\multicolumn{7}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\ \hline
	    \textbf{Name} & \textbf{Year} & \textbf{Tor interoperability} & \textbf{Public implementation} & \textbf{Used in practice} & \textbf{Attack resistance} & \textbf{Provides anonymity} \\ \hline
		\endhead
		\hline \multicolumn{7}{r}{\textit{Continued on next page}} \\
		\endfoot
		\hline
		\endlastfoot

		Gnutella & 2000 & \xmark & \checkmark & \checkmark & \xmark & \xmark & \xmark & \xmark \\ \hline
		Freenet & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \checkmark \\ \hline
		Tapestry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
		Pastry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
		Anonymizing peer-to-peer proxy (AP3) & 2004 & \xmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
		Tarzan & 2002 & \xmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
		MorphMix & 2002 & \xmark & \checkmark & \xmark & \checkmark & \xmark & ? & \checkmark \\ \hline
		I2P & 2003 & \xmark & \checkmark & \checkmark & ? & ? & ? & \checkmark\\ \hline
		Salsa & 2006 & \xmark & \xmark & \xmark & \checkmark & \checkmark & ? & \checkmark \\ \hline
		Tribler & 2008 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
		Torsk & 2009 & \checkmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
		NISAN & 2009 & \xmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
    \end{longtable}
\end{center}