### 2014-03-21 table.tex

\begin{tabular}[p]{| l | l | p{2.5cm} | p{2.5cm} | p{2.5cm} | p{2.5cm} | p{2.5cm} | p{2.5cm} | p{2.5cm} |}
	\hline
	\textbf{Name} & \textbf{Year} & \textbf{Tor interoperability} & \textbf{Public implementation} & \textbf{Used in \newline practice} & \textbf{(D)DoS protection} & \textbf{Sybil attack protection} & \textbf{MITM \newline protection} & \textbf{Provides \newline anonymity} \\ \hline

	Gnutella & 2000 & \xmark & \checkmark & \checkmark & \xmark & \xmark & \xmark & \xmark \\ \hline
	Freenet & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \checkmark \\ \hline
	Tapestry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	Pastry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	AP3 & 2004 & \xmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
	Tarzan & 2002 & \xmark & \xmark & \xmark & ? & \checkmark & ? & \checkmark \\ \hline
	MorphMix & 2002 & \xmark & \checkmark & \xmark & \checkmark & \xmark & ? & \checkmark \\ \hline
	I2P & 2003 & \xmark & \checkmark & \checkmark & \checkmark & \checkmark & ? & \checkmark\\ \hline
	Salsa & 2006 & \xmark & \xmark & \xmark & \checkmark & \checkmark & ? & \checkmark \\ \hline
	Tribler & 2008 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	Torsk & 2009 & \checkmark & \xmark & \xmark & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
	NISAN & 2009 & \xmark & \xmark & \xmark & ? & ? & \checkmark & \checkmark \\ \hline		
\end{tabular}