### 2014-03-24 tbl-alternativeinternet.tex
And the actual table

\begin{tabular}[p]{| r | l | p{3cm} | p{3cm} | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
	\hline
	\textbf{\#} & \textbf{Name} & \textbf{Language} & \textbf{Age} & \textbf{Last activity} & \textbf{LOC} & \textbf{Commits} & \textbf{Contributors} \\ \hline

	1 & ownCloud & PHP & 6 year(s) & 2014-03-15 & 1,297 K & 34,391 commit(s) & 392 contributor(s) \\ \hline
	2 & Freenet & Java & 13 year(s) & 2014-03-15 & 442 K & 32,009 commit(s) & 183 contributor(s) \\ \hline
	3 & Tor & C & 12 year(s) & 2014-03-13 & 329 K & 29,200 commit(s) & 184 contributor(s) \\ \hline
	4 & GNUnet & C & 8 year(s) & 2014-03-13 & 427 K & 21,398 commit(s) & 37 contributor(s) \\ \hline
	5 & StatusNet & PHP & 6 year(s) & 2014-03-14 & 230 K & 14,877 commit(s) & 92 contributor(s) \\ \hline
	6 & Diaspora* & Ruby & 3 year(s) & 2014-03-14 & 51 K & 14,202 commit(s) & 368 contributor(s) \\ \hline
	7 & SlapOS & Python & 8 year(s) & 2014-03-15 & 588 K & 14,046 commit(s) & 93 contributor(s) \\ \hline
	8 &Tahoe-LAFS & Python & 7 year(s) & 2014-03-13 & 158 K & 11,565 commit(s) & 54 contributor(s) \\ \hline
	9 & Tribler & Python & 8 year(s) & 2014-03-15 & 148 K & 11,521 commit(s) & 48 contributor(s) \\ \hline
	10 & Lorea & PHP & 6 year(s) & 2014-03-12 & 873 K & 11,066 commit(s) & 96 contributor(s) \\ \hline
\end{tabular}