### 2014-03-26 tbl-currentsystems.tex
Modified table to use vertically oriented column headers

\begin{tabular}[p]{| l | l | l | l | l | l | l | l | l |}
	\hline
	\parbox[t]{2mm}{\textbf{Name}} & \parbox[t]{2mm}{\textbf{Year}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{Tor interoperability}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{Public implementation}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{Used in practice}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{(D)DoS protection}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{Sybil attack protection}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{MITM protection}}} & \parbox[t]{2mm}{\rotatebox[origin=l]{90}{\textbf{Unlinkability}}} \\ \hline

	Gnutella & 2000 & \xmark & \checkmark & \checkmark & \xmark & \xmark & \xmark & \xmark \\ \hline
	Freenet & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \checkmark \\ \hline
	Tapestry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	Pastry & 2001 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	Tarzan & 2002 & \xmark & \xmark & \xmark & ? & \checkmark & ? & \checkmark \\ \hline
	MorphMix & 2002 & \xmark & \checkmark & \xmark & \checkmark & \xmark & ? & \checkmark \\ \hline
	I2P & 2003 & \xmark & \checkmark & \checkmark & \checkmark & \checkmark & ? & \checkmark\\ \hline
	AP3 & 2004 & \xmark & \xmark & \xmark & ? & ? & ? & \checkmark \\ \hline
	Salsa & 2006 & \xmark & \xmark & \xmark & \checkmark & \checkmark & ? & \checkmark \\ \hline
	Tribler & 2008 & \xmark & \checkmark & \checkmark & \checkmark & \xmark & \xmark & \xmark \\ \hline
	Torsk & 2009 & \checkmark & \xmark & \xmark & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
	NISAN & 2009 & \xmark & \xmark & \xmark & ? & ? & \checkmark & \checkmark \\ \hline		
\end{tabular}